Improving \textsl{Gaia} parallaxes with a data-driven model \\ of the color--magnitude diagram

### 2017-03-20 paperDistance/paper.tex

Here we build and use just such a data-driven model. In particular,
we use all of the \gaia\ \tgas\ data (\citealt{tgas};
these data were generated by the methodology of \citealt{michalik15}), matched to the
\tmass\ photometry (\citealt{skrutskie06}), and in the footprint of
the \panstarrs\ Dust Map (\citealt{green16}), to make a model of the
noise-free (or low-noise) color--magnitude diagram (\cmd) of stars.
We use an Empirical-Bayes approach known as Extreme Deconvolution (XD;
\citealt{bovy11}).
This method deconvolves heteroskedastic data to derive an
estimate of the noise-free or low-noise distribution that \emph{would
  have} been observed with far better data.
This method has been used in astrophysics
previously to model the velocity distribution of stars in the Solar
Neighborhood (\citealt{hogg05}; \citealt{bovy09}) and to
perform quasar target selection
in \project{Sloan Digital Sky Survey} data (\citealt{xdqso}; \citealt{xdqsoz}).
Its advantage over other deconvolution methods is that it takes as input,
and handles in a principled way, heteroskedastic data.
Its principle disadvantage for the present purposes is that it requires
or implies a strictly Gaussian noise model.
In the below, this requires us to transform the \cmd\ to a space in which
the noise is approximately Gaussian.

>>>>

Here we build and use just such a data-driven model. In particular,
we use all of the \gaia\ \tgas\ data (\citealt{tgas};
these data were generated by the methodology of \citealt{michalik15}), matched to the
\tmass\ photometry (\citealt{skrutskie06}), and in the footprint of
the \panstarrs\ Dust Map (\citealt{green15}), to make a model of the
noise-free (or low-noise) color--magnitude diagram (\cmd) of stars.
We use an Empirical-Bayes approach known as Extreme Deconvolution (XD;
\citealt{bovy11}).
This method deconvolves heteroskedastic data to derive an
estimate of the noise-free or low-noise distribution that \emph{would
  have} been observed with far better data.
This method has been used in astrophysics
previously to model the velocity distribution of stars in the Solar
Neighborhood (\citealt{hogg05}; \citealt{bovy09}) and to
perform quasar target selection
in \project{Sloan Digital Sky Survey} data (\citealt{xdqso}; \citealt{xdqsoz}).
Its advantage over other deconvolution methods is that it takes as input,
and handles in a principled way, heteroskedastic data.
Its principle disadvantage for the present purposes is that it requires
or implies a strictly Gaussian noise model.
In the below, this requires us to transform the \cmd\ to a space in which
the noise is approximately Gaussian.
